---
title: "R Notebook"
output: html_notebook
---


```{r}
setwd('F:/resume_project/k_means_clustering/KMeans/Dataset')

```

```{r}
require(readr)
data<-read_csv('data.csv')
```
```{r}
str(data)
```


```{r}
head(data)
```


```{r}
colSums(is.na(data))
```


```{r}
data$Joined<-NULL
data$`Loaned From`<-NULL
data$`Contract Valid Until`<-NULL
data$LS<-NULL
data$ST<-NULL
data$RS<-NULL
data$LW<-NULL
data$LF<-NULL
data$CF<-NULL
data$RF<-NULL
data$LAM<-NULL
data$CAM<-NULL
data$RAM<-NULL
data$LM<-NULL
data$LCM<-NULL
data$CM<-NULL
data$RCM<-NULL
data$RM<-NULL
data$LWB<-NULL
data$LDM<-NULL
data$CDM<-NULL
data$RDM<-NULL
data$RWB<-NULL
data$LB<-NULL
data$LCB<-NULL
data$CB<-NULL
data$RCB<-NULL
data$RB<-NULL
data$`Release Clause`<-NULL
data$Club<-NULL
data$Position<-NULL


```

#there may be the case that players are new that they didnt play any match that's why there #many attributes are missing but we cant omit all these  players as they are in large #number so better to remove all those variables 


```{r}
data<-na.omit(data)
```
#as there are only 48 missing values so we can ommit them as it doesnot lead to loss of much information and its better to remove then to put false information


```{r}
colSums(is.na(data))
```


```{r}
require(dplyr)

data<-mutate_if(data,is.character,tolower)
data<-distinct(data)


```



```{r}
Overall_gp<-group_by(data,ID)%>% summarise(Overall_gp=mean(Overall))
```



```{r}
potential<-group_by(data,ID)%>% summarise(potential=mean(Potential))
```



```{r}
aggression<-group_by(data,ID)%>% summarise(aggression=mean(Aggression))
```



```{r}
agility<-group_by(data,ID)%>% summarise(agility=mean(Agility))
```


```{r}
Age<-group_by(data,ID)%>% summarise(Age=mean(Age))
```


```{r}
dribbling<-group_by(data,ID)%>% summarise(dribbling=mean(Dribbling))
```


```{r}
volleys<-group_by(data,ID)%>% summarise(volleys=median(Volleys))
```


```{r}
acc<-group_by(data,ID)%>% summarise(acc=median(Acceleration))
```

```{r}
GKPos<-group_by(data,ID)%>% summarise(GKPos=mean(GKPositioning))
```


```{r}
Tackle<-group_by(data,ID)%>% summarise(Tackle=mean(StandingTackle))
```

```{r}
new_data<- merge(potential,Overall_gp, by = 'ID')

new_data <- merge(new_data, aggression, by = 'ID')


new_data <- merge(new_data, agility, by = 'ID')

new_data <- merge(new_data, acc, by = 'ID')

new_data <- merge(new_data,Age, by = 'ID')
new_data <- merge(new_data,GKPos , by = 'ID')
new_data <- merge(new_data,dribbling , by = 'ID')
new_data <- merge(new_data,Tackle , by = 'ID')
new_data <- merge(new_data,volleys , by = 'ID')
new_data <- merge(new_data, potential, by = 'ID')

```



```{r}
new_data[,-1] <- sapply(new_data[,-1], scale)

```


#missing values
```{r}
colSums(is.na(new_data))
```


Outlier treatment

aggression

```{r}
plot(quantile(new_data$aggression, seq(0, 1, 0.01)),na.rm=T)
quantile(new_data$aggression, seq(0, 1, 0.01))

```

seems like no as such outliers




overall_gp

```{r}
plot(quantile(new_data$Overall_gp, seq(0, 1, 0.01)))
quantile(new_data$Overall_gp, seq(0, 1, 0.01))

```

```{r}
new_data$Overall_gp[new_data$Overall_gp >  2.43409363    ] <-  2.43409363   
new_data$Overall_gp[new_data$Overall_gp < -2.27550756  ] <- -2.27550756  

```



#potential

```{r}
plot(quantile(new_data$potential.x, seq(0, 1, 0.01)))
quantile(new_data$potential.x, seq(0, 1, 0.01))

```

```{r}
new_data$potential.x[new_data$potential.x < -2.06038531 ] <- -2.06038531
new_data$potential.x[new_data$potential.x > 2.54911377 ] <- 2.54911377

```


agility

```{r}
plot(quantile(new_data$agility, seq(0, 1, 0.01)))
quantile(new_data$agility, seq(0, 1, 0.01))

```

```{r}
new_data$agility[new_data$agility < -2.70810237 ] <- -2.70810237

```


#Age
```{r}
plot(quantile(new_data$Age, seq(0, 1, 0.01)))
quantile(new_data$Age, seq(0, 1, 0.01))

```

```{r}
new_data$Age[new_data$Age > 2.417961719 ] <- 2.417961719

```


#acc
```{r}
plot(quantile(new_data$acc, seq(0, 1, 0.01)))
quantile(new_data$acc, seq(0, 1, 0.01))

```

```{r}
new_data$acc[new_data$acc < -2.972905929  ] <- -2.972905929

```

#GKPos
```{r}
plot(quantile(new_data$GKPos, seq(0, 1, 0.01)))
quantile(new_data$GKPos, seq(0, 1, 0.01))

```

```{r}
new_data$GKPos[new_data$GKPos > 1.776679304 ] <- 1.776679304
new_data$GKPos[new_data$GKPos < -1.795654722 ] <- -1.795654722
```

# dribbling
```{r}
plot(quantile(new_data$dribbling, seq(0, 1, 0.01)))
quantile(new_data$dribbling, seq(0, 1, 0.01))

```

```{r}
new_data$dribbling[new_data$dribbling < -2.84449585 ] <- -2.84449585

new_data$dribbling[new_data$dribbling > 1.87100469 ] <- 1.87100469

```

#Tackle
```{r}
plot(quantile(new_data$Tackle, seq(0, 1, 0.01)))
quantile(new_data$Tackle, seq(0, 1, 0.01))

```

```{r}
new_data$Tackle[new_data$Tackle > 1.569149383 ] <- 1.569149383

```

#volleys
```{r}
plot(quantile(new_data$volleys, seq(0, 1, 0.01)))
quantile(new_data$volleys, seq(0, 1, 0.01))

```

```{r}
new_data$volleys[new_data$volleys < -1.87695480 ] <- -1.87695480

```

#potential.y
```{r}
plot(quantile(new_data$potential.y, seq(0, 1, 0.01)))
quantile(new_data$potential.y, seq(0, 1, 0.01))

```

```{r}
new_data$potential.y[new_data$potential.y > 2.54911377 ] <-  2.54911377

new_data$potential.y[new_data$potential.y < -2.06038531 ] <-  -2.06038531

```



outliers treated

Implementing K-Means algorithm

Finding the optimal value of K

```{r}
r_sq <- rep(0, 30)
r_sq

for (number in 1:30) { 
          cluster_k <- kmeans(new_data[,-1], centers = number, nstart = 20, iter.max = 50)
          r_sq[number] <- cluster_k$betweenss/cluster_k$totss }

```


```{r}
plot(r_sq, type = 'l')

points(r_sq, col = 'blue')

```



```{r}

# cluster 3 

clusters_3 <- kmeans(new_data[,c(2:12)],centers = 3,iter.max = 50, nstart = 20)
  
new_data$cluster3 <- as.factor(clusters_3$cluster)
clusters_3$centers
table(clusters_3$cluster)




```

```{r}
# cluster 4 

clusters_4 <- kmeans(new_data[,c(2:12)],centers = 4,iter.max = 50, nstart = 20)

new_data$cluster4 <- as.factor(clusters_4$cluster)
clusters_4$centers
table(clusters_4$cluster)
```



```{r}

# cluster 5 

clusters_5 <- kmeans(new_data[,c(2:12)],centers = 5,iter.max = 50, nstart = 20)

new_data$cluster5 <- as.factor(clusters_5$cluster)
clusters_5$centers
table(clusters_5$cluster)



```

```{r}
# try fviz_cluster for visualization
require(cluster)
require(factoextra)


new_data$cluster3 <- as.numeric(new_data$cluster3)
new_data$cluster4 <- as.numeric(new_data$cluster4)
new_data$cluster5 <- as.numeric(new_data$cluster5)

fviz_cluster(clusters_3,data = new_data)
fviz_cluster(clusters_4,data = new_data)
fviz_cluster(clusters_5,data = new_data)


```
from graph it is clear that cluster 3 is performing much better than cluster4 and cluster5

```{r}
# See clusters behavior

require(dplyr)
df <- group_by(new_data, cluster3) %>% summarise(avg_overall_gp = mean(Overall_gp),
                                      avg_pot.x = mean(potential.x),
                                      avg_pot.y = mean(potential.y),
                                      avg_agg = mean(aggression),
                                      avg_agility = mean(agility),
                                      avg_acc = mean(acc),
                                      avg_age = mean(Age),
                                      avg_GKPos= mean(GKPos),
                                      avg_dribb = mean(dribbling),
                                      avg_Tackle = mean(Tackle),
                                      avg_vol = mean(volleys))

df

```

```{r}

# May be by visualizing we get the clear idea also
require(ggplot2)
ggplot(df, aes(x = avg_overall_gp, y = avg_pot.x,
               size = avg_agg,color=avg_agility)) + geom_point() +geom_label(aes(label=cluster3))


```

```{r}
# See clusters behavior

require(dplyr)
df_4 <- group_by(new_data, cluster4) %>%  summarise(avg_overall_gp = mean(Overall_gp),
                                      avg_pot.x = mean(potential.x),
                                      avg_pot.y = mean(potential.y),
                                      avg_agg = mean(aggression),
                                      avg_agility = mean(agility),
                                      avg_acc = mean(acc),
                                      avg_age = mean(Age),
                                      avg_GKPos= mean(GKPos),
                                      avg_dribb = mean(dribbling),
                                      avg_Tackle = mean(Tackle),
                                      avg_vol = mean(volleys))
df_4

```

```{r}

# May be by visualizing we get the clear idea also
require(ggplot2)
ggplot(df_4, aes(x = avg_overall_gp, y = avg_pot.y,
               size = avg_agg,color=avg_agility)) + geom_point() +geom_label(aes(label=cluster4))


```


```{r}
# See clusters behavior

require(dplyr)
df_5 <- group_by(new_data, cluster5) %>%  summarise(avg_overall_gp = mean(Overall_gp),
                                      avg_pot.x = mean(potential.x),
                                      avg_pot.y = mean(potential.y),
                                      avg_agg = mean(aggression),
                                      avg_agility = mean(agility),
                                      avg_acc = mean(acc),
                                      avg_age = mean(Age),
                                      avg_GKPos= mean(GKPos),
                                      avg_dribb = mean(dribbling),
                                      avg_Tackle = mean(Tackle),
                                      avg_vol = mean(volleys))

df_5

```

```{r}

# May be by visualizing we get the clear idea also
require(ggplot2)
ggplot(df_5, aes(x = avg_overall_gp, y = avg_pot.y,
               size = avg_agg,color=avg_agility)) + geom_point() +geom_label(aes(label=cluster5))


```



#from cluster 3 we can have clear cut idea about three different type of players ,players #who belongs to cluster 1 have low avg_overall_gp and avg_potential but high avg_aggression #and agility average is also low
 
 # players from cluster 2 have high avg_overall_gp,avg_aggression and avg_potential but low
 #avg_agility


# and from cluster 3 player ahve high avg_potential and low avg_overall_gp as compared to cluster 1 have high avg_aggression and low avg_agility


#so I choose 3 clusters


```{r}

require(dplyr)
df <- group_by(new_data, cluster3) %>% summarise(avg_overall_gp = mean(Overall_gp),
                                      avg_pot.x = mean(potential.x),
                                      avg_pot.y = mean(potential.y),
                                      avg_agg = mean(aggression),
                                      avg_agility = mean(agility),
                                      avg_acc = mean(acc),
                                      avg_age = mean(Age),
                                      avg_GKPos= mean(GKPos),
                                      avg_dribb = mean(dribbling),
                                      avg_Tackle = mean(Tackle),
                                      avg_vol = mean(volleys))

df

```